### 2019-02-01 thesis_report/methods.tex
Added to methods and theory

\section{Physical picture}
	\comment{We consider a two-dimensional system with zigzag.}
	We consider a Josephson junction (JJ)(Fig.~\ref{fig:setup}) consisting of a two dimensional strip of semiconductor, with superconductors on both sides.
	We modulate the shape of the normal region, which can be either zigzag as depicted, or a more smooth sinusoidal like shape.
	Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
	We model the system with a BdG Hamiltonian Eq.~\eqref{eq:hamiltonian}; the normal part has a linear Rashba spin-orbit coupling term characterized by $\alpha$ and a Zeeman field with $E_z=\frac{1}{2}mu_B g B_x$.
	The superconductor has a superconducting coupling term $\Delta$ with a superconducting phase $\exp{\pm i \phi/2}$.
	Both parts have a kinetic term and chemical potential $\mu$ (counted from the bottom spin-orbit band). 
	For the zigzag geometry, we first consider a sawtooth like pattern where $z_x$ is periodicity, $z_y$ the amplitude and $W$ the width of the junction.

@@

	\comment{The symmetry class of the system is D.}
	The non-zigzag system is in the symmetry class BDI \cite{pientka2017topological}.
	Using the software package Qsymm\cite{varjas2018qsymm}, we find that the symmetry class of our system to be $\mathcal{D}$ because we break the chiral/time-reversal resulting from the mirror symmetric junction.

	\begin{small}
	\begin{align}
	    H = \left(\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu + \frac{\meff \alpha^2}{2 \hbar^2} \right)\tau_z+
	        E_\text{z} \sigma_x+
	        \alpha \left( \ky \sigma_x - \kx \sigma_y \right) \tau_z +
	        \Delta \tau_y
	\end{align}
	\label{eq:hamiltonian}
	\end{small}

@@

	\comment{We use kwant to make systems and extract the eigenstates}
	We use a tight-binding model, implemented using Kwant, to simulate various properties of the system.
	The model consists of a two-dimensional square lattice of sites.
	Each site has two orbitals: spin and electron-hole.
	There are four different types of sites (and thus four different Hamiltonians): the top superconductor, the bottom superconductor, the normal region, and a normal region with a variable potential offset, forming a barrier to simulated an imperfect NS interface.
	A schematic of such kwant systems is depicted in Fig.~\ref{fig:kwant_system}.
	We can also make the system infinite in for example the $x$ direction by connecting the sites on the left border with those on the right border with a $\kx$ dependent hopping.

@@

	\comment{We extract the gap by either taking the lowest energy state for an infinite system, or the difference between the lowest two eigenstates.}
	After the system is specified, we can extract the total Hamiltonian of the system, allowing analysis of the eigenstates.
	For example, calculating the bandstructure of a system is performed by taking a translationally invariant system and calculating the eigenvalues for different momenta.

>>>>

\section{Physical picture}
	\comment{We consider a two-dimensional system with zigzag.}
	We consider a Josephson junction (JJ)(Fig.~\ref{fig:setup}) consisting of a two dimensional strip of semiconductor, with superconductors on both sides.
	We modulate the shape of the normal region, which can be either zigzag as depicted, or a more smooth sinusoidal like shape.
	Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
	We model the system with a BdG Hamiltonian, Eqs.~\eqref{eq:hamiltonian_n}-~\eqref{eq:hamiltonian_s}  ; the normal part has a linear Rashba spin-orbit coupling term characterized by $\alpha$ and a Zeeman field with $E_z=\frac{1}{2}mu_B g B_x$.
	The superconductor has a superconducting coupling term $\Delta$ with a superconducting phase $\exp{\pm i \phi/2}$.
	Both parts have a kinetic term and chemical potential $\mu$ (counted from the bottom spin-orbit band). 
	For the zigzag geometry, we first consider a sawtooth like pattern where $z_x$ is periodicity, $z_y$ the amplitude and $W$ the width of the junction.

@@

	\begin{align}
	    H_\text{N} &= \left(\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu + \frac{\meff \alpha^2}{2 \hbar^2} \right) \left( \sigi \otimes \tauz \right)+
	        E_\text{z} \left(\sigx \otimes \tauz \right) +
	        \alpha \left( \ky \sigx - \kx \sigma_y \right) \otimes \tauz \label{eq:hamiltonian_n} \\
	    H_\text{SC} &= \left(\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu \right) \left(\sigi \otimes \tauz \right) +
	        \Delta \left(\sigi \otimes \tau_{\pm \phi /2}\right)
	\label{eq:hamiltonian_s}
	\end{align}
	where,
	\begin{equation}
	    \tau_{\pm \phi /2} = \cos\left( \pm \phi /2 \right) \taux + \sin\left( \pm \phi /2 \right) \tauy \nonumber
	\end{equation}

\subsection{Symmetry class of system}
	\comment{The symmetry class of the system is D.}
	The non-zigzag system is in the symmetry class BDI \cite{pientka2017topological}, unless the additional symmetries are broken due to for example a difference in superconducting gap magnitude between the two superconductors.
	Using the software package Qsymm\cite{varjas2018qsymm}, we find that the symmetry class of our system to be $\mathcal{D}$, because we break this chiral/time-reversal symmetry resulting from the mirror symmetric junction.

@@

	\comment{We use kwant to make systems and extract the eigenstates}
	We use a tight-binding model, implemented using Kwant, to simulate various properties of the system.
	The model consists of a two-dimensional square lattice of sites.
	Each site has two orbitals: spin and electron-hole.
	There are four different types of sites (and thus four different Hamiltonians): the top superconductor, the bottom superconductor, the normal region, and a normal region with a variable potential offset, forming a barrier to simulate an imperfect NS interface.
	A schematic of such kwant systems is depicted in Fig.~\ref{fig:kwant_system}.

	\comment{The model hamiltonian is converted into the tight binding form by substitution}
	In order to turn the set of Hamiltonians, Eqs.\eqref{eq:hamiltonian_n} and \eqref{eq:hamiltonian_s} into a tight binding model, we write the equations into differential form by substituting $k_\alpha \rightarrow - i \partial_\alpha$, $\alpha \in \left\{ x, y \right\}$, and thereafter discretize on a rectangular grid.
	In practice, this is implemented using the Kwant discretizer module.

@@

	\comment{We extract the gap by either taking the lowest energy state for an infinite system, or the difference between the lowest two eigenstates.}
	After the system is specified, we can extract the total Hamiltonian of the system, allowing analysis of the eigenstates.
	For example, calculating the bandstructure of a system is performed by taking a translationally invariant system and calculating the eigenvalues for different momenta.
	The system is made translationally invariant, in for example the $x$ direction, by connecting the sites on the left border with those on the right border with a $\kx$ dependent hopping.

### 2019-02-01 thesis_report/theory.tex
Simple terms figure, updated theory and introduction
Added to methods and theory

\chapter{Theory}
The topological quantum computer uses pairs of Majorana particles as its unit of computation (qubit).
In order to host Majorana modes in a system, one needs a material which has properties from a superconducting material, as well as from a semi-conducting material.
As this material does not seem to exist on its own, one creates a heterogeneous system.
I simple terms, this consists of a slab of superconductor and a strip of semiconductor.
By interfacing the two, one creates a region where all the necessary ingredients are present to realize Majorana's.

@@

\section{Normal region}
	Two important ingredients required for realizing Majorana fermions are strong spin-orbit coupling, and a strong interaction with the applied magnetic field.
	It is the normal conductor which brings these two physical effects to the mix.
	Note that the Hamiltonian of the normal region is the same for electrons and holes, except for the sign.
	The Hamiltonian terms provided in this section are relevant to the electron states.

@@

		For the purpose of creating Majorana's, we are interested in two main aspects of superconductivity: the energy gap surrounding the Fermi energy (due to the depletion of states), and the electron-hole (/particle-hole) symmetry.
		We are thus not interested in the paired up electrons (Cooper pairs), and only wish to model the electronic (unpaired) modes.
		This can be achieved by modeling superconductivity with the Boguliubov de Gennes Hamiltonian:

		\begin{equation}
		\mathbf{H}_\text{BdG} = \begin{bmatrix} H & \Delta \\ \Delta^\dagger & H \end{bmatrix}
		\end{equation}

		The block matrix restricts the structure of the system in the electron-hole basis, and ensures particle-hole symmetry.
		For $\Delta \neq 0$ one indeed induces a identically sized gap in the spectrum.

>>>>

\chapter{Theoretical Background}
The topological quantum computer uses pairs of Majorana particles as its unit of computation (qubit).
In order to host Majorana modes in a system, one needs a material which has properties from a superconducting material, as well as from a semi-conducting material.
As this material does not seem to exist on its own, one creates a heterogeneous system.
I simple terms, this consists of a slab of superconductor and a strip of semiconductor.
By interfacing the two, one creates a region where all the necessary ingredients are present to realize Majorana's.

@@

\section{Normal region}
	Two important ingredients required for realizing Majorana fermions are strong spin-orbit coupling, and a strong interaction with the applied magnetic field.
	It is the semiconductor which brings these two physical effects to the mix.
	Note that the Hamiltonian of the normal region is the same for electrons and holes, except for the sign.
	The Hamiltonian terms provided in this section are relevant to the electron states.
	In figure \ref{fig:ham_simple_terms}, the effect of the various terms in the Hamiltonian on the dispersion relation is shown.

	\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.95\columnwidth]{figures/simple_terms}
	\caption{Energy dispersion of electrons with different physical effects (spin orbit and magnetic field).
	The dashed line in the figure with both spin orbit coupling and zeeman field corresponds to the case when magnetic field is not aligned with the spin orbit direction.}
	\label{fig:ham_simple_terms}
	\end{figure}

	\subsection{Kinetic term}
		We rewrite the kinetic term in the Schrodinger equation into it's momentum form by noting $- i \partial_x \rightarrow \kx$:
		\begin{equation}
			H_\text{kin} = - \frac{\hbar^2}{2 \meff} \nabla^2 = \frac{\hbar^2}{2 \meff} \left( \kx^2 + \ky^2 \right)
		\end{equation}
		The effective mass is a way of capturing the periodic potential of the crystal lattice into a single term.
		Its value differs per material.

@@

	\subsection{Superconducting Hamiltonian}
		For the purpose of creating Majorana's, we are interested in two main aspects of superconductivity: the energy gap surrounding the Fermi energy (due to the depletion of states), and the electron-hole (/particle-hole) symmetry.
		We are thus not interested in the paired up electrons (Cooper pairs), and only wish to model the electronic (unpaired) modes.
		This can be achieved by modeling superconductivity with the Boguliubov de Gennes Hamiltonian, which ensures particle-hole symmetry:

		\begin{equation}
		\mathbf{H}_\text{BdG} = \begin{bmatrix} H & \Delta \\ \Delta^\dagger & -H \end{bmatrix}
		\label{eq:bdg}
		\end{equation}

		Note that Eq.\eqref{eq:bdg} describes a block matrix: both $H$ and $\Delta$ can be matrices themselves, where the degrees of freedom correspond to for example spin, or discretized positions.
		Using this basis, one is able to describe a heterogeneous system, containing both superconducting and normal regions.
		As an example, we write the Hamiltonian of a heterogeneous 1D system, consisting of a superconductor extending from $x = -\infty$ to $0$, and a normal system from  $x = 0$ to $\infty$:

		\begin{equation}
			\mathbf{H} =  \Theta \left(-x\right) \begin{bmatrix} H_\text{SC, kin} & \Delta \\ \Delta^\dagger & -H_\text{SC, kin} \end{bmatrix} + \Theta \left(x\right)  \left( \tau_z \otimes H_\text{normal} \right) 
			\end{equation}
		where, 
		\begin{equation}
			\Theta(x) = \begin{cases} 1 & x \geq 0 \\ 0 & x < 0 \end{cases}
		\end{equation}
